---
title: "Conformity Model"
author: "Dan Conroy-Beam"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Model Overview 
 
### Purpose

The purpose of this model is to explore the emergence of fads as a function of varying levels of conformity and anticonformity.

Agents will choose between three options (grow a beard, grow a mustache, or be clean-shaven) based on one of two conformity strategies. Some agents will be "conformists": these agents will adopt facial hair styles in proportion to how popular they are in their social group. Other agents will be "hipsters": these agents will adopt facial hair styles in proportion to how unpopular they are in their social group.
 
### Agents

A population of agents will be generated. Each agent will have a strategy: hipster or conformist. The population size, proportion of hipsters, probability of changing hairstyle, and social group size will be determined by manipulable parameters.

Agents will be assigned to random social groups; social groups will be non-overlapping such that not all agents within a particular agent's social group will mutually know one another.
 
### Life Cycle
 
1. Assess
2. Switch
 
#### Assess

Each agent will first assess the facial hair of all other agents in their social group. The agent will then determine the proportion of agents who have beards, mustaches, and are clean shaven.
 
#### Switch

After all agents have assessed the population, each agent will switch their facial hair based on their conformity strategy.  

The probability of agents changing their facial hair each month will depend on a manipulable parameter.  

"Conformists" will choose a random facial hair style with probability directly proportional to the proportion of agents in their social group with that facial hair style.  

"Hipsters" will choose a random facial hair style with probability inversely proportional to the proportion of agents in their social group with that facial hair style.  

In this way, conformists will tend toward wearing whatever is most popular whereas hipsters will tend toward wearing whatever is least popular
 
### Analysis

The model will plot the proportion of agents wearing each facial hair type as a function of time.


## Model



### Packages

```{r packages}

library(ggplot2)

```



### Parameters

```{r parameters}

#Population Size#
#The number of agents to generate
popSize<-100



#Proportion of Hipsters#
#The proportion of agents who adopt an anticonformity strategy
propHipsters<-.01



#Probability of Hairstyle Change#
#The probability agents will change their facial hairstyle each month
pChange<-.33



#Social Group Size#
#The size of each agent's social group
groupSize<-30



#Months#
#Number of months to simulate
months<-1000

```



### Functions

```{r functions}

#Agent Generation#
#A function to generate agents
agentGenerate<-function(n,p){
  
  #Generate agent conformity strategies
  strategy<-c(rep("conformist",(1-p)*n),rep("hipster",p*n))
  
  #Generate agent facial hair styles
  facialHair<-sample(c("beard","clean","stash"),n,replace=T)
  
  #Put the agents together
  agents<-data.frame(strategy,facialHair)
  
  #Output the agents
  return(agents)
}



#Socialize#
#A function to generate agents social networks
socialize<-function(n,g){
  
  #Generate a blank matrix to represent the social network
  network<-matrix(0,n,n)
  
  #Loop through agents...
  for(a in 1:n){
    
    #Assign each agent a random number of agents whom they "know"
    network[a,sample((1:n)[-a],g)]<-1
    
  }
  
  #Output the agents' social network
  return(network)
  
}

```



### Life Cycle

```{r lifeCycle}

#Generate the agents
agents<-agentGenerate(popSize,propHipsters)

#Generate the agents' social network
network<-socialize(popSize,groupSize)

#Make a blank dataframe to store model results
results<-data.frame("time"=1:months,
           "style"=rep(c("beard","clean","stash"),each=months),
           "prop"=NA
           )

#Loop through months
for(m in 1:months){
  
  
  #Assess#
  
  #Generate a blank matrix to store popularity of each facial hair type in each agent's social network
  propStyle<-matrix(NA,popSize,3)
  
  #Loop through agents...
  for(a in 1:popSize){
    
    #Compute the proportion of other agents in agent a's social network wearing each facial hair style
    propStyle[a,1]<-mean(agents$facialHair[network[a,]==1]=="beard")
    propStyle[a,2]<-mean(agents$facialHair[network[a,]==1]=="clean")
    propStyle[a,3]<-mean(agents$facialHair[network[a,]==1]=="stash")

  }
  
  
  
  #Switch#
  
  #Loop through agents...
  for(a in 1:popSize){
    
    #If the agent decides to change their facial hair...
    if(rbinom(1,1,pChange)){
      
      if(agents$strategy[a]=="conformist"){
      
        #Have the agent pick a hairstyle in proportion to popularity in their network
        agents$facialHair[a] <- sample(c("beard","clean","stash"),1,
               prob=propStyle[a,])
    
      } else {
        
        #Have the agent pick a hairstyle in proportion to unpopularity in their network
        agents$facialHair[a] <- sample(c("beard","clean","stash"),1,
               prob=1-propStyle[a,])
        
      }
    }
  }
  
  results$prop[results$time==m & results$style=="beard"] <- mean(agents$facialHair=="beard")
  results$prop[results$time==m & results$style=="clean"] <- mean(agents$facialHair=="clean")
  results$prop[results$time==m & results$style=="stash"] <- mean(agents$facialHair=="stash")
  
}


```



### Analysis

```{r analysis}

p1 <- ggplot(data=results,
             aes(x=time,
                 y=prop,
                 fill=style)
             )+
  labs(x="Time (in Months)",
       y="Popularity of Facial Hair Style",
       fill="Facial Hair Style")+
  scale_fill_manual(labels=c("Beard","Clean-shaven","Moustache"),
                             values=c("red","yellow","forestgreen"))+
  geom_area()

p1

```

